### 2011-02-10 paper/results.tex
changes
slight change to red triangle description
HN results

\acmFiggggg{explore_dim_1}{explore_dim_2}{explore_dim_3}{explore_dim_4}{explore_dim_5}{1}{.53}{Fitness
  mean and standard deviation when each parameter dimension is varied
  independently.  The red triangle is the intersection of the slices
  along each dimension and is common to all plots.}

@@

The results for the parameterized gaits are shown in
\figref{std_error} and \tabref{results}.  As a group, these strategies
produced results not significantly better than the randomly-generated
gaits, however, all methods managed to explore the space enough to
significantly improve on the previous hand coded gait in at least one
of the three runs.  No single strategy consistently outperformed the
others: for the first trial Linear Regression produced the fastest
gait at 27.58 body lengths per minute, for the second a random gait
actually won with 17.26 body lengths / minute, and for the third trial
the Nelder-Mead simplex method attained the fastest gait at 14.83 body
lengths / minute.

@@

%%%%%%% NEW TABLE
\begin{table}
\begin{center}
\begin{tabular}{|r|c|c|c||c|}
\hline
                                         & Average & Std. Dev. \\
\hline                                    
\hline                                    
Previous hand-coded gait                 & 5.16   &   --     \\
\hline
Random search                            & 9.40   &   6.83   \\
\hline
Uniform Random Hill Climbing             & 7.83   &   4.56   \\
\hline
Gaussian Random Hill Climbing            & 10.03  &   6.00   \\
\hline
Policy Gradient Descent                  & 6.32   &   7.39   \\
\hline
Nelder-Mead simplex                      & 12.32  &   3.35   \\
\hline
Linear Regression                        & 14.01  &  12.88   \\
\hline
Evolutionary Neural Network              &        &          \\
(HyperNEAT)                              & 29.26  &   6.37   \\
\hline
\end{tabular}
\caption{The best gaits found for each starting vector and algorithm,
  in body lengths per minute.}
\tablabel{results}
\end{center}
\end{table}

@@

\subsection{Evolved Neural Network Strategy}

The evolved neural network strategy used was HyperNEAT. HyperNEAT
produced the fastest gaits by far. Linear regression produced a gait
in the first trial (27.58 body lengths/minute) that was on par with
the best gaits generated by HyperNEAT, but no other maximizing or
predictive strategy produced gaits close to those of HyperNEAT. Linear
regression was the best performing maximizing or predictive strategy,
but it resulted in an average best gait of only 14.01 body
lengths/minute. HyperNEAT, on the other hand, resulted in an average
best gait of 29.26 body lengths/minute. One issue encountered with
HyperNEAT was that it generated gaits that were rougher on the robot
than gaits generated by the other strategies. The gaits sometimes
caused the motor servos to overheat, which caused the motor to behave
differently than expected. Since an overheated motor produces a
different walk that expected, a test was performed after each run to
make sure all motors were working correctly. If a motor failed, the
distance walked would be halved to penalize walks that were harmful to
the robot.

\edit{take this out?}

We have done complete runs of 3 different initial parameter vectors
for random search, uniform random hill climbing, Gaussian random hill
climbing, policy gradient descent, Nelder-Mead simplex, and linear
regression.  We also evaluated the SVM regression and HyperNEAT
methods, but these methods were more experimental, and thus we do not
yet have the same volume of data for these runs.  We developed several
gaits that were about 4 times faster than the original hand-coded
gait.  Results are shown in \tabref{results}.

\subsection{Discussion}

% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.

Because only three trials were tested with each algorithm and no
algorithm consistently outperformed the others, there is a large
standard error for each algorithm, as show in \figref{std_error}. For
this reason, it is unclear if any algorithm is superior to another in
this application. More trials with each algorithm would be necessary
to reach a definitive ranking of the performance of the algorithms.

However, each algorithm discovered at least one gait of over 10 body
lengths/minute, including random. For this reason, we speculate that
the motion model is at least as critical as the learning algorithm
used.

\acmFig{neat_110115_211410_00000_002_filt}{1}{Caption here...???}
\acmFig{neat_110115_211410_00000_002_filt_zoom}{1}{Caption here...???}

>>>>

\acmFiggggg{explore_dim_1}{explore_dim_2}{explore_dim_3}{explore_dim_4}{explore_dim_5}{1}{.53}{Fitness
  mean and standard deviation when each parameter dimension is varied
  independently.  The red triangle in each plot represents the same point in the 5-dimensional parameter space.}

@@

The results for the parameterized gaits are shown in
\figref{std_error} and \tabref{results}.  As a group, these strategies
produced results not significantly better than the randomly-generated
gaits, however, all methods managed to explore the space enough to
significantly improve on the previous hand coded gait in at least one
of the three runs.  No single strategy consistently outperformed the
others: for the first trial Linear Regression produced the fastest
gait at 27.58 body lengths/minute, for the second a random gait
actually won with 17.26 body lengths/minute, and for the third trial
the Nelder-Mead simplex method attained the fastest gait at 14.83 body
lengths/minute.

@@

%%%%%%% NEW TABLE
\begin{table}
\begin{center}
\begin{tabular}{|r|c|c|c||c|}
\hline
                                         & Average & Std. Dev. \\
\hline                                    
\hline                                    
Previous hand-coded gait                 & 5.16   &   --     \\
\hline
Random search                            & 9.40   &   6.83   \\
\hline
Uniform Random Hill Climbing             & 7.83   &   4.56   \\
\hline
Gaussian Random Hill Climbing            & 10.03  &   6.00   \\
\hline
Policy Gradient Descent                  & 6.32   &   7.39   \\
\hline
Nelder-Mead simplex                      & 12.32  &   3.35   \\
\hline
Linear Regression                        & 14.01  &  12.88   \\
\hline
Evolved Neural Network              &        &          \\
(HyperNEAT)                              & 29.26  &   6.37   \\
\hline
\end{tabular}
\caption{The best gaits found for each starting vector and algorithm,
  in body lengths per minute.}
\tablabel{results}
\end{center}
\end{table}

@@

\subsection{HyperNEAT Gaits}

The results for the gaits produced and evolved by HyperNEAT are shown
in \figref{hnResults} and \tabref{results}.  Overall the evolved gaits
were the fastest by far, beating all the parameterized models when
comparing either the average or the best gaits of each.

\acmFig{hnResults}{1}{Average fitness ($\pm $ standard error) of
  generation champion (highest performing individual among the
  population of nine individuals) for each generation of the three
  HyperNEAT runs.}

%\acmFig{neat_110115_211410_00000_002_filt}{1}{Caption here...???}

\acmFig{neat_110115_211410_00000_002_filt_zoom}{1}{Example of one
  particular high-performance gait produced by HyperNEAT showing
  commands for each of nine motors.  Note the complexity of the motion
  pattern. Such patterns were not possible with the parameterized
  SineModel5, nor would they likely be contrived by a human designing
  a different low-dimensional parameterized motion model.}
